Canopy leaf area of an evergreen Eucalyptus woodland does not respond to elevated

### 2015-05-08 manuscript.Rmd
data analysis description improved

```{r echo=FALSE, message=FALSE}
library(knitr)
load("cache/lai_workspace.RData")
source("load.R")

@@

# Number of digits in printing of numbers
options(scipen = 1, digits = 2)

@@

\
In ecosystems where water availability limits productivity, we could expect eC~a~ to further affect *L* [@Donohue2009], since eC~a~ reduces stomatal conductance, and hence water use per unit of leaf area [@Warren2010] leading to increased soil water availability. These ‘water savings’ can be expected to lead to an increase in *L* in water-limited ecosystems, consistent with a tight coupling of *L* with precipitation across gradients [@Ma2013; @Jin2013; @Hatton2008]. In support of this prediction, Donohue et al. [-@Donohue2013] attributed the observed global increase in *L* in arid and semi-arid areas over the last 30 years to rising C~a~. Based on the same water-savings principle, we predict that *L* should increase under eC~a~ in water-limited forests.

@@

\
To overcome violations in the assumptions underlying Eq. 2, we calibrated *L_\tau* against cumulative litter production during a drought period in 2013 (indicated in Fig. 1). During this period (14 July 2013 – 12 Nov 2013), no leaf growth occurred but *L* decreased steadily (and thus $\tau_d$ increased steadily, see Fig. 1). During the drought, only 42 mm precipitation was recorded, which was the second driest July-October period recorded at a nearby (<2km) weather station since 1881 (compared to a long-term average of 188 mm). A total of n=11 days with sufficiently diffuse conditions to estimate $\tau_d$  were available during this period. Cumulative *L* leaf litter production ($P_L$, expressed in units of *L*) should match the observed decrease in *L* estimated from canopy transmittance. We calibrated Eq. 3,

@@

where $\tau_p$ canopy transmission estimated from canopy photos. We assumed that the extinction coefficient *k* = 0.5, following Macfarlane et al. [-@Macfarlane2007] recommendation for eucalypt forests. Without further calibration, the two estimates were reasonably well correlated, though on average $L_\tau$ was `r 100*(with(flatcan_byring, mean(LAI.PAR.mean / LAI))-1)` higher than $L_p$ (Fig. SI2, $R^2$=`r glance(lmphoto)$r.squared`, RMSE=`r glance(lmphoto)$sigma`) (Fig. S2). We also estimated the extinction coefficient that minimized the difference between the two estimates of *L* (Fig. S2), giving *k* = `r kopt`. These comparisons are encouraging because the photos measure *L* near the zenith (+/- 15°), while $L_\tau$ is integrated over the sky hemisphere (Eq. 3), weighted towards the zenith.

## Data analysis
We used generalized additive models (gam) for all analyses of *L* and $\tau_d$ timeseries [@Wood2006], using the *mgcv* package in R [@RCT2015]. For all time-series, we used a base of `r .gambase` when fitting the gams, which resulted in biologically realistic smoothed *L* dynamics. In all fits, we used the FACE ring as a random effect, except when we analysed average *L* across all rings (Fig. 7). We estimated the derivative of the fitted gam, including confidence intervals for the derivative, following Simpson [@Simpson2014]. For comparison between *L* from photos and PAR transmittance, we used simple linear regression. To test for effects of C~a~ treatment on $L_\tau$ and litter production, we used linear-mixed effects models with FACE ring as the random effect, and C~a~ treatment and date as a factor.

@@

@@

# Figure captions {.unnumbered}

@@

@@

@@

@@

@@

@@

@@

@@

@@

>>>>

```{r echo=FALSE, message=FALSE}
library(knitr)

load("cache/lai_workspace.RData")
source("load.R")

@@

# Number of digits in printing of numbers
options(scipen = 1, digits = 3)

@@

\
In ecosystems where water availability limits productivity, we could expect eC~a~ to further affect *L* [@Donohue2009], since eC~a~ reduces stomatal conductance, and hence water use per unit of leaf area [@Warren2010] leading to increased soil water availability. These ‘water savings’ can be expected to lead to an increase in *L* in water-limited ecosystems, consistent with a tight coupling of *L* with precipitation across gradients [@Ma2013; @Jin2013; @Ellis2008]. In support of this prediction, Donohue et al. [-@Donohue2013] attributed the observed global increase in *L* in arid and semi-arid areas over the last 30 years to rising C~a~. Based on the same water-savings principle, we predict that *L* should increase under eC~a~ in water-limited forests.

@@

\
To overcome violations in the assumptions underlying Eq. 2, we calibrated $L_\tau$ against cumulative litter production during a drought period in 2013 (indicated in Fig. 1). During this period (14 July 2013 – 12 Nov 2013), no leaf growth occurred but *L* decreased steadily (and thus $\tau_d$ increased steadily, see Fig. 1). During the drought, only 42 mm precipitation was recorded, which was the second driest July-October period recorded at a nearby (<2km) weather station since 1881 (compared to a long-term average of 188 mm). A total of n=11 days with sufficiently diffuse conditions to estimate $\tau_d$  were available during this period. Cumulative *L* leaf litter production ($P_L$, expressed in units of *L*) should match the observed decrease in *L* estimated from canopy transmittance. We calibrated Eq. 3,

@@

where $\tau_p$ canopy transmission estimated from canopy photos. We assumed that the extinction coefficient *k* = 0.5, following Macfarlane et al. [-@Macfarlane2007] recommendation for eucalypt forests. Without further calibration, the two estimates were reasonably well correlated, though on average $L_\tau$ was `r 100*(with(flatcan_byring, mean(LAI.PAR.mean / LAI))-1)`% higher than $L_p$ (Fig. SI2, $R^2$=`r glance(lmphoto)$r.squared`, RMSE=`r glance(lmphoto)$sigma`) (Fig. S2). We also estimated the extinction coefficient that minimized the difference between the two estimates of *L* (Fig. S2), giving *k* = `r kopt`. These comparisons are encouraging because the photos measure *L* near the zenith (+/- 15°), while $L_\tau$ is integrated over the sky hemisphere (Eq. 3), weighted towards the zenith.

## Data analysis
We used generalized additive models (GAM) for all analyses of *L* and $\tau_d$ timeseries [@Wood2006], using the *mgcv* package in R [@RCT2015]. A GAM is a semi-parametric fitting method that allows the estimation of smoothed trends in the data, including confidence intervals. In all fitted GAMs, we used a cubic regression spline. For the smoothed term in the model (days since start of experiment), we used up to `r .gambase` degrees of freedom, which resulted in biologically realistic smoothed *L* dynamics. Random effects were included to account for the grouping of data within FACE rings, except when we analysed average *L* across all rings (Fig. 7). For Fig. 7, we estimated the derivative of the fitted GAM, including confidence intervals for the derivative, following Simpson [-@Simpson2014; see also @Curtis2014].

\
For comparison between *L* from photos and PAR transmittance, we used simple linear regression. To test for effects of C~a~ treatment on $L_\tau$ and litter production, we used linear-mixed effects models with FACE ring as the random effect, and C~a~ treatment and date as factor variables.

@@

\
The entire analysis presented in this paper will be made available as a *git* repository (http://www.github.com/remkoduursma/XXXXX), published on figshare (doi:XXXX). The repository includes the processed data needed to run the analyses and figures (*raw data needs to be shared too, have to figure out how though!*).

@@

# Figures {.unnumbered}

```{r}
options(scipen = 1, digits = 5)
```

```{r fig.width=8, fig.height=6, echo=FALSE, message=FALSE}
figure1(subset(facegap_cloudy_byring,Date < maxdate), ramp)
```

@@

```{r fig.width=4, fig.height=4, echo=FALSE, message=FALSE}
figure2(ba)
```

@@

```{r fig.width=8, fig.height=4, echo=FALSE, message=FALSE}
figure3(subset(facegap_cloudy_byring, Date < maxdate))
```

@@

```{r fig.width=6, fig.height=4, echo=FALSE, message=FALSE, echo=FALSE, message=FALSE}
figure4(dLAIlitter)
```

@@

```{r fig.width=8, fig.height=4, echo=FALSE, message=FALSE}
figure5(dLAIlitter)
```

@@

```{r fig.width=8, fig.height=4, echo=FALSE, message=FALSE}
figure6(ba)
```

@@

```{r fig.width=10, fig.height=8, echo=FALSE, message=FALSE}
figure7(subset(facegap_cloudy_byring, Date < maxdate),
               subset(facesoilwater, Date < maxdate),
               subset(faceraindaily, Date < maxdate),
               subset(airt, Date < maxdate))
```

@@

```{r fig.width=5, fig.height=8, echo=FALSE, message=FALSE}
figureSI1(litring, subset(facegap_cloudy_byring, Date < maxdate))
```

@@

```{r fig.width=8, fig.height=4, echo=FALSE, message=FALSE}
figureSI2(flatcan_byring)
```

@@

```{r fig.width=8, fig.height=5, echo=FALSE, message=FALSE}
figureSI3(subset(facegap_cloudy_byring, Date < maxdate), 
          subset(facegap_all_byring, Date < maxdate))
```
